---
title: "Missing, delayed, and old: A status review of ESA recovery plans<br><span class='note'>Jacob Malcom, Defenders of Wildlife<br>Sept. 2016</span>"
output: 
  html_document:
    highlight: tango
    fig_caption: yes
    fig_width: 7
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, message = FALSE}
library(ecosscraper)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(knitr)
library(lubridate)
library(plotly)
library(recovery.plan.overview)
library(stringr)
library(viridis)

data("recovery_plan_table")
data("species_info_table")
data("ecos_all_spp")
data("nmfs_man")
```

# Reference

This vignette accompanies a manuscript we are submitting to _PeerJ Preprints_ and _PeerJ_. The preprint can be found [here](#).

# Abstract

Recovery planning is an essential part of implementing the U.S. Endangered Species Act (ESA), but conservationists and government agencies recognize there are problems with the process. A common perception is that too many species lack recovery plans, these plans take too long to write, and plans are rarely updated to include new information. Using data from all U.S. domestic and transboundary ESA-listed species--those species we expect should have recovery plans--we quantify basic characteristics of ESA recovery planning. We show that ~1/4 of these listed species lack recovery plans; the average recovery plan has taken >5 years to be developed after listing; and half of recovery plans are 19 or more years old. These results are not unexpected given dwindling budgets and more species to protect, but they underscore the need for systematic improvements to recovery planning. We discuss how many of the shortcomings we identify here can be ameliorated, at least in part, by transitioning to modern, web-based recovery plans. 

# Data prep

```{r prep_data, message = FALSE}
###########################################################################
# load and prep the data
#
# First, fix some errors from ECOS
recovery_plan_table$Date <- ifelse(recovery_plan_table$Title == "Final Recovery Plan for Johnson's Seagrass",
                                   "2002-09-15",
                                   recovery_plan_table$Date)
recovery_plan_table$Date <- ifelse(recovery_plan_table$Title == "White Abalone Recovery Plan",
                                   "2008-10-10",
                                   recovery_plan_table$Date)
recovery_plan_table$Date <- as_date(recovery_plan_table$Date)

# Add manually curated data from NMFS
combo <- rbind(recovery_plan_table, nmfs_man)

# # Now do table joins and filtering
# combo <- dplyr::left_join(with_new,
#                           species_info_table,
#                           by = "Species")
# names(combo)[1] <- "Rec_Plan_Date"
# names(combo)[4] <- "Plan_Status"
# names(combo)[8] <- "Date_Cur_Status"
# combo$Date_Cur_Status <- as_date(combo$Date_Cur_Status)
# combo$Cur_Status_To_Plan <- as.numeric(combo$Rec_Plan_Date - combo$Date_Cur_Status)
# combo <- distinct(combo, Title, Species, `Where Listed`, .keep_all = TRUE)
# combo$Status <- sub(combo$Status, 
#                     pattern = 'displayListingStatus\\(\\"', 
#                     replacement = '')
# combo$Status <- sub(combo$Status, 
#                     pattern = '\\"\\)', 
#                     replacement = '')
# combo <- filter(combo, Status == "Threatened" | Status == "Endangered")

# Next, the join with the 'big' ECOS species table to get original listing dates
names(combo)[1] <- "Rec_Plan_Date"
names(combo)[4] <- "Plan_Status"
recplan_spp <- left_join(combo, ecos_all_spp, by = c("Species" = "Scientific_Name"))
recplan_spp <- filter(recplan_spp, US_Foreign %in% c("US", "US/Foreign"))
recplan_spp <- filter(recplan_spp, 
                      Federal_Listing_Status %in% c("Endangered", "Threatened"))
recplan_spp <- select(recplan_spp, Rec_Plan_Date, Title, Plan_Status, 
                      Species, Common_Name, Federal_Listing_Status, Lead_Region, 
                      First_Listed, Species_Group, US_Foreign, Where_Listed)
recplan_spp <- distinct(recplan_spp, Species, Where_Listed, .keep_all = TRUE)
recplan_spp$Elapsed <- as.numeric(recplan_spp$Rec_Plan_Date - recplan_spp$First_Listed)
recplan_spp$Elapsed_Years <- recplan_spp$Elapsed / 365

# Looks like I'm missing records for ~50 species compared to what the BoxScore
# provides (http://ecos.fws.gov/ecp0/reports/box-score-report)...wait, no, the
# footnote on that page says "active plans," so both draft and final.
final_plans <- filter(recplan_spp, Plan_Status == "Final")
with_plan <- filter(recplan_spp, 
                    recplan_spp$Plan_Status != "Outline",
                    recplan_spp$Plan_Status != "Conservation Strategy",
                    recplan_spp$Plan_Status != "Exempt")

###############
# Also need the complement of the previous, a data.frame of listed species,
# with NAs for listed species lacking plans. Note that the 1593 I get here is 
# the same as what BoxScore indicates today.
listed <- filter(ecos_all_spp, 
                 Federal_Listing_Status %in% c("Endangered", "Threatened"))
listed <- filter(listed, US_Foreign %in% c("US", "US/Foreign"))

listed <- left_join(listed, combo, by = c("Scientific_Name" = "Species"))
listed <- distinct(listed, Scientific_Name, Where_Listed, .keep_all = TRUE)
listed$Year_Listed <- year(listed$First_Listed)
listed$Year_Plan <- year(listed$Rec_Plan_Date)

# We also want to see what's happened with time-to-plan through time,
# and need to calculate a "minimum" date at which final plans might exist.
# We replace any NA or date for species without a final or revised plan
# with 15 Sep 2016 as a placeholder. Also need to NA green sea turtle because
# the plan was done in 1999, but the species was split into 6 DPSs in 2016
listed$Min_Date_noFinal <- rep(NA, length(listed$Scientific_Name))
for(i in 1:length(listed$Min_Date_noFinal)) {
  if(is.na(listed$Rec_Plan_Date[i])) {
    listed$Min_Date_noFinal[i] <- as.Date("2016-09-15", origin = "1970-01-01")
  } else if(grepl(listed$Scientific_Name[i], pattern = "Chelonia mydas")) {
    listed$Min_Date_noFinal[i] <- NA
  } else if(grepl(listed$Plan_Status[i], pattern = "Final|Revision")) {
    listed$Min_Date_noFinal[i] <- as.Date(listed$Rec_Plan_Date[i])
  } else {
    listed$Min_Date_noFinal[i] <- as.Date("2016-09-15", origin = "1970-01-01")
  }
}
listed$Min_Date_noFinal <- as.Date(listed$Min_Date_noFinal, origin = "1970-01-01")
listed$Age_Min_Age <- as.numeric(listed$Min_Date_noFinal - listed$First_Listed)
# hist(listed$Age_Min_Age)

no_plan <- filter(listed, is.na(listed$Year_Plan))
```

# Analyses

## Summary statistics

```{r summary_stats}
delta_tmp <- c(n = length(final_plans$Elapsed_Years),
               min = round(min(final_plans$Elapsed_Years, na.rm = TRUE), 1),
               median = round(median(final_plans$Elapsed_Years, na.rm = TRUE), 1),
               mean = round(mean(final_plans$Elapsed_Years, na.rm = TRUE), 1),
               max = round(max(final_plans$Elapsed_Years, na.rm = TRUE)))
listd_tmp <- c(n = length(final_plans$First_Listed ),
               min = as.character(min(final_plans$First_Listed, na.rm = TRUE)),
               median = as.character(median(final_plans$First_Listed, na.rm = TRUE)),
               mean = as.character(mean(final_plans$First_Listed, na.rm = TRUE)),
               max = as.character(max(final_plans$First_Listed, na.rm = TRUE)))
pland_tmp <- c(n = length(final_plans$Rec_Plan_Date),
               min = as.character(min(final_plans$Rec_Plan_Date, na.rm = TRUE)),
               median = as.character(median(final_plans$Rec_Plan_Date, na.rm = TRUE)),
               mean = as.character(mean(final_plans$Rec_Plan_Date, na.rm = TRUE)),
               max = as.character(max(final_plans$Rec_Plan_Date, na.rm = TRUE)))
to_show_df <- rbind(listd_tmp, pland_tmp, delta_tmp)
row.names(to_show_df) <- c("Listed Date", "Plan Date", "Years Elapsed")

kable(to_show_df,
      caption = "Summary statistics of time between listing and final recovery plan. Min, median, mean, and max are given in days.",
      digits = 1,
      align = "c")
```

# Figures

## Figure 1a: Number of species listed, with/out plans

```{r figure_1a, message = FALSE}
years <- seq(min(listed$Year_Listed, na.rm = TRUE),
             max(listed$Year_Listed, na.rm = TRUE))
n_listed <- unlist(
              lapply(years, 
                FUN = function(x) {
                  sum(listed$Year_Listed <= x,  na.rm = TRUE)
                }
              )
            )
n_plans <- unlist(
             lapply(years, 
               FUN = function(x) {
                 sum(listed$Year_Plan <= x & 
                     (listed$Plan_Status == "Final" |
                       grepl(listed$Plan_Status, pattern = "Revision")), 
                                               na.rm = TRUE)
               }
             )
           )
n_list_df <- data.frame(years, n_listed, n_plans)
n_list_df$n_no_plan <- n_list_df$n_listed - n_list_df$n_plans

ggplot(n_list_df, aes(y = n_listed, x = years)) +
  geom_line() +
  geom_line(aes(y = n_no_plan, x = years),
            colour = "gray70",
            size = 2) +
  geom_line(aes(y = n_plans, x = years), 
            colour = "gray70",
            linetype = "dashed") +
  labs(x = "", y = "Count") +
  theme_hc() + 
  theme(text = element_text(family = "Open Sans"))
```

## Figure 1b: Plans/species with plans through time

```{r figure_1b, message = FALSE}
n_listed <- unlist(lapply(years, 
                          function(x) sum(listed$Year_Listed == x, na.rm = TRUE)))
n_w_plans <- unlist(lapply(years, 
                         function(x) {
                           sub <- filter(listed, Year_Listed == x)
                           sum(!is.na(sub$Year_Plan))
                         }))
n_species <- unlist(lapply(years, 
                         function(x) {
                           sub <- filter(listed, Year_Plan == x)
                           length(unique(sub$Scientific_Name))
                         }))
n_plans <- unlist(lapply(years, 
                         function(x) {
                           sub <- filter(listed, Year_Plan == x)
                           length(unique(sub$Title))
                         }))
n_list_df <- data.frame(years, n_listed, n_w_plans, n_species, n_plans)
n_list_df$pct_w_plan <- n_list_df$n_w_plans / n_list_df$n_listed

ggplot(n_list_df, aes(y = n_plans, x = years)) +
  geom_line(colour = "gray70") +
  geom_line(aes(y = n_species, x = years)) +
  labs(x = "Year",
       y = "Count") +
  theme_hc() +
  theme(text = element_text(family = "Open Sans"))
```

Number of species without a recovery plan in September 2016: 
**`r n_list_df$n_no_plan[length(n_list_df$n_no_plan)]`**

## Figure 2

```{r figure_2, message = FALSE}
ggplot(n_list_df, aes(y = pct_w_plan, x = years)) +
  geom_point(size = 3, alpha = 0.5) +
  geom_smooth(level = 0.9, alpha = 0, colour = "black") +
  scale_y_continuous(limits = c(0,1)) +
  labs(x = "Year Listed",
       y = "Proportion with plans") +
  theme_hc() + 
  theme(text = element_text(family = "Open Sans"))
```


## Figure 3a: Times-to-plan (histogram)

```{r figure_3a, message = FALSE}
qplot(final_plans$Elapsed_Years, geom = "histogram") +
      labs(x = "Years between date listed and recovery plan date",
           y = "Number of species") +
      theme_hc() +
      theme(plot.title=element_text(hjust = 0, size = 18),
            text = element_text(family = "Open Sans"))
```

## Figure 3b: Times-to-plan (percentile)

```{r figure_3b, message = FALSE}
duration <- sort(final_plans$Elapsed_Years, decreasing = FALSE)
order <- seq(1:length(duration))
percentile <- order / max(order)
pctile <- data.frame(duration, order, percentile, stringsAsFactors = FALSE)

ggplot(pctile, aes(y = duration, x = percentile)) +
  geom_line() +
  labs(x = "Percentile",
       y = "Time-to-recovery plan (years)\n") +
  theme_hc() + 
  theme(text = element_text(family = "Open Sans"))
# ggplotly(gg) %>% layout(font = list(family = "Open Sans"))
```

## Figure 4: Times-to-plan through time

```{r figure_4, message = FALSE}
tmp <- final_plans
tmp$Year_Listed <- year(tmp$First_Listed)
tmp$Year_Plan <- year(tmp$Rec_Plan_Date)
years <- seq(min(tmp$Year_Listed, na.rm = TRUE),
             max(tmp$Year_Listed, na.rm = TRUE))

cur_fx <- function(df, x, fn) {
  sub_df <- filter(df, Year_Listed == x)
  if(dim(sub_df)[1] == 0) return(NA)
  switch(fn,
         mean = mean(sub_df$Elapsed_Years, na.rm = TRUE),
         min = min(sub_df$Elapsed_Years, na.rm = TRUE),
         max = max(sub_df$Elapsed_Years, na.rm = TRUE)
  )
}

# This is to calculate a potential "minimum time-to-plan" for species 
# that do not have final plans or final plans in revision. However, 
# the results aren't particularly informative and are not included in
# the paper or the figure below.
get_potential <- function(df, x) {
  sub_df <- filter(df, Year_Listed == x)
  if(dim(sub_df)[1] == 0) return(NA)
  mean = mean(sub_df$Age_Min_Age, na.rm = TRUE)
  return(mean / 365)
}

n_plans <- unlist(lapply(years, function(x) {length(filter(tmp, Year_Plan == x)[[1]])} ))
tm_mean <- unlist(lapply(years, cur_fx, df = tmp, fn = "mean"))
tm_min <- unlist(lapply(years, cur_fx, df = tmp, fn = "min"))
tm_max <- unlist(lapply(years, cur_fx, df = tmp, fn = "max"))
potential_age <- unlist(lapply(years, get_potential, df = listed))
dat <- data.frame(years, tm_mean, tm_min, tm_max, potential_age, n_plans)
names(dat) <- c("Year", "Mean", "Min", "Max", "Potential", "N_plans")

ggplot(dat, aes(y = Mean, x = Year)) +
  geom_line() +
  geom_line(aes(y = Min, x = Year), colour = "gray70") +
  geom_line(aes(y = Max, x = Year), colour = "gray70") +
  # geom_line(aes(y = Potential, x = Year), colour = "red") +
  labs(x = "Year of Listing",
       y = "Time-to-plan (years)\n") +
  theme_hc() + 
  theme(text = element_text(family = "Open Sans"))
```

# Plan ages

```{r plan_age, echo = FALSE, warning = FALSE, message = FALSE, fig.cap="Click figure to enlarge"}
with_plan$Current_Age <- as.numeric(Sys.Date() - with_plan$Rec_Plan_Date) / 365

qplot(with_plan$Current_Age, geom = "histogram") +
      labs(x = "Plan age (years)",
           y = "Number of species") +
      ggtitle("Age of recovery plans") +
      theme_hc() +
      theme(plot.title=element_text(hjust = 0, size = 18))
```

With a median age of `r round(median(with_plan$Current_Age, na.rm = TRUE), 2)` years, most recovery plans are quite old - if plans were people, they could vote. The distribution reflects the pattern of plans-through-time even without considering the 135 plans that have been or are being revised. As with the time-to-plan, we can use a dynamic percentile plot to quickly find out the proportion of species' plans that are less than X years old:

```{r spp_pctile, echo = FALSE, warning = FALSE, message = FALSE, fig.cap="Click figure to enlarge"}
age <- sort(with_plan$Current_Age, decreasing = FALSE)
order <- seq(1:length(age))
percentile <- order / max(order)
pctile <- data.frame(age, order, percentile, stringsAsFactors = FALSE)

gg <- ggplot(pctile, aes(y = age, x = percentile)) +
        geom_line() +
        ggtitle("Age of species' plans") +
        labs(x = "Percentile",
             y = "Plan age (years)\n") +
        theme_hc() + 
        theme(text = element_text(family = "Open Sans"))

ggplotly(gg) %>% layout(font = list(family = "Open Sans"))
```

Only 16% of species have plans that are < 10 years old; 10% of species have plans that are >30 years old. These numbers highlight a dire need for some way to improve how recovery plans are updated. Many species are part of multi-species plans, and so have the same age for the individual plans. We can plot the ages of plans, rather than of species' plan ages:

```{r plan_pctile, echo = FALSE, warning = FALSE, message = FALSE, fig.cap="Click figure to enlarge"}
tmp <- distinct(with_plan, Title, .keep_all = TRUE)
age <- sort(tmp$Current_Age, decreasing = FALSE)
order <- seq(1:length(age))
percentile <- order / max(order)
pctile <- data.frame(age, order, percentile, stringsAsFactors = FALSE)

gg <- ggplot(pctile, aes(y = age, x = percentile)) +
        geom_line() +
        ggtitle("Ages of recovery plans") +
        labs(x = "Percentile",
             y = "Plan age (years)\n") +
        theme_hc() + 
        theme(text = element_text(family = "Open Sans"))

ggplotly(gg) %>% layout(font = list(family = "Open Sans"))
```

This view of the data suggests a slightly worse picture. The median age of _plans_ is 22 years and 17% of plans are > 30 years old; the difference on the high end is because more older plans are single-species than the average-age plan.

# Plans by region and agency

Because of the high degree of independence among FWS regions and differences between FWS and NMFS, it may be instructive to break down recovery planning along those lines. _Which regions / agency have finalized the highest numbers and proportions of plans?_

```{r tab_region, echo = FALSE, warning = FALSE}
mat <- as.matrix(table(listed$Lead_Region, is.na(listed$Year_Plan)))
res <- data.frame(Region_Agency = row.names(mat),
                  With_Plans = mat[, 1],
                  Without_Plans = mat[, 2],
                  Prop_Plans = mat[, 1] / (mat[,1] + mat[,2]))
row.names(res) <- NULL
kable(res, 
      caption = "Distribution of plans among FWS regions and NMFS.", 
      align = "c",
      digits = 3)
```

How does the variation in time-to-plan break down by region/agency?

```{r by_region, echo = FALSE, warning = FALSE, fig.cap="Click figure to enlarge"}
ggplot(final_plans, aes(y = Elapsed_Years, x = Lead_Region)) +
  geom_violin(fill = "gray", alpha = 0.5, colour = NA) +
  geom_boxplot(colour = "gray20", fill = NA) +
  labs(x = "FWS Region / NMFS", y = "Time-to-Plan (years)") +
  ggtitle("Time to recovery plan by region / agency") +
  theme_hc()
```

Time for NMFS is both high (median) and highly variable relative to FWS regions. Regions 4 and 8, which encompass many species, do not take as long recovery planning as I would have thought.  

# Discussion

Recovery planning was incorporated in the 1978 amendments to the ESA because of the realization that planning should improve conservation outcomes. Significant progress has been made improving the quality of recovery plans: contemporary plans are far more detailed and science-based than many older plans. But the handful of species that benefit from "modern" recovery plans are overshadowed by the shortcomings identified here.

_First_, 1/4 of ESA-listed species currently lack a recovery plan, even a draft that is publicly available. How can anyone other than specialists know what to do to help - or at least avoid harming - these listed species? If someone discovers an important tool for the conservation of one of these species, how can they know whether anyone else knows? Recovery outlines are a good starting point, but the data indicate only 64 of the 428 species lacking recovery plans have a recovery outline. Draft recovery plans that are readily available are the next step up, but there are only 34 of those on ECOS. Filling the gap with at least some well-informed guidance for the ~350 species lacking any type of plan should be a high priority.

_Second_, and as recognized by the review panel for NMFS's recovery program review, the time between listing an providing a plan is too long. While the median time is "only" 5.3 years, that number is pulled down by two factors. Nearly 120 species have plans that were finalized before those species were listed; removing those raises the median time to 5.9 years. More drastically, those 450 species that currently lack a recovery plan are excluded from the analysis and their addition even today would certainly increase the median time-to-plan. The NMFS review panel report provides a substantial number of recommendations that can likely reduce the time to first recovery planning documents and "final" plans. Building on the preceding paragraph, we would encourage the Services to, at a minimum, get recovery outlines in place for all listed species. Doing so will at least jump-start recovery planning for the overlooked species, and we expect will lead to reducing the time-to-plan in the coming years.

_Third_, at a median of 19 years and 10% of plans 30 years or older, many completed plans are showing their age. Not only is it likely that science has learned more about these species and their conservation, the threats to species may have changed significantly over these extended timeframes. For example, the indigo snake [recovery plan](https://ecos.fws.gov/docs/recovery_plan/820422.pdf) was finalized in 1982; poaching may have been a significant threat at that time, but habitat destruction is clearly the leading threat today. Recognizing that formal recovery plan updates, as traditionally practiced, are time-consuming and expensive (e.g., just in _Federal Register_ notices), the Services need to transition to a new and more dynamic recovery planning framework.

We recognize that the Services understand through experience some of the challenges that this analysis has shown. For example, the Service has been developing their [Recovery Enhancement Vision (REV)](http://www.nmfs.noaa.gov/pr/recovery/plans/usfws_recovery_enhancement_vision_overview_april_2016_draft.pdf) for a number of years, which should help improve several aspects of recovery planning. (And NMFS expressed their interest in the REV model in their response to the recent review.) We think that the [dynamic, web-based recovery plan](https://defend-esc-dev.org/dynamic-recovery/) frameworks that we are developing can be particularly useful in implementing REVs and similar updates to recovery planning. For example, recovery plans for the 450 species lacking plans - and hundreds of species that will be listed in the coming years - can start as recovery outlines in the web-based framework. Through regular updates the outlines can be fleshed out as draft plans, and after public review, "finalized." These online plans would be continually updated rather than being treated as fixed documents. Rather than a future analysis finding the average recovery plan hasn't been updated in 19 years, we think it is completely feasible that the dynamic plans could have an average "last update" of 19 days or weeks instead. Such a future can only help improve conservation outcomes of imperiled species.

# Appendix A _Data cleaning_

While the code for all of the plots and tables can be easily viewed by examining the Rmarkdown document for this vignette, we provide the data cleaning (commented out) here for ease of viewing from the HTML document.

```{r est_orig_list}
# combo <- dplyr::left_join(recovery_plan_table,
#                           species_info_table,
#                           by = "Species")
# names(combo)[1] <- "Rec_Plan_Date"
# combo$Rec_Plan_Date <- as_date(combo$Rec_Plan_Date)
# names(combo)[4] <- "Plan_Status"
# names(combo)[8] <- "Date_Cur_Status"
# combo$Date_Cur_Status <- as_date(combo$Date_Cur_Status)
# combo$Cur_Status_To_Plan <- as.numeric(combo$Rec_Plan_Date - combo$Date_Cur_Status)
# combo <- distinct(combo, Species, `Where Listed`, .keep_all = TRUE)
# combo$Status <- sub(combo$Status, 
#                     pattern = 'displayListingStatus\\(\\"', 
#                     replacement = '')
# combo$Status <- sub(combo$Status, 
#                     pattern = '\\"\\)', 
#                     replacement = '')
# combo <- filter(combo, Status == "Threatened" | Status == "Endangered")
# 
# # Next, the join with the 'big' ECOS species table to get original listing dates
# recplan_spp <- left_join(combo, ecos_all_spp, by = c("Species" = "Scientific_Name"))
# recplan_spp <- filter(recplan_spp, US_Foreign %in% c("US", "US/Foreign"))
# recplan_spp <- filter(recplan_spp, 
#                       Federal_Listing_Status %in% c("Endangered", "Threatened"))
# recplan_spp <- select(recplan_spp, Rec_Plan_Date, Title, Plan_Status, 
#                       Species, Status, Date_Cur_Status, Lead_Region, 
#                       Cur_Status_To_Plan, First_Listed, Species_Group,
#                       US_Foreign, Where_Listed)
# recplan_spp <- distinct(recplan_spp, Species, Where_Listed, .keep_all = TRUE)
# recplan_spp$Elapsed <- as.numeric(recplan_spp$Rec_Plan_Date - recplan_spp$First_Listed)
# recplan_spp$Elapsed_Years <- recplan_spp$Elapsed / 365
# 
# # Looks like I'm missing records for ~50 species compared to what the BoxScore
# # provides (http://ecos.fws.gov/ecp0/reports/box-score-report)...wait, no, the
# # footnote on that page says "active plans," so both draft and final.
# final_plans <- filter(recplan_spp, recplan_spp$Plan_Status == "Final")
# with_plan <- filter(recplan_spp, 
#                     recplan_spp$Plan_Status != "Outline",
#                     recplan_spp$Plan_Status != "Conservation Strategy",
#                     recplan_spp$Plan_Status != "Exempt")
# 
# ###############
# # Also need the complement of the previous, a data.frame of listed species,
# # with NAs for listed species lacking plans. Note that the 1593 I get here is 
# # the same as what BoxScore indicates today.
# listed <- filter(ecos_all_spp, 
#                  Federal_Listing_Status %in% c("Endangered", "Threatened"))
# listed <- filter(listed, US_Foreign %in% c("US", "US/Foreign"))
# 
# listed <- left_join(listed, combo, by = c("Scientific_Name" = "Species"))
# listed <- distinct(listed, Scientific_Name, Where_Listed, .keep_all = TRUE)
# listed <- select(listed, -`Plan Action Status`, -Doc_Link, -Status, -`Lead Region`,
#                   -`Where Listed`)
# listed$Year_Listed <- year(listed$First_Listed)
# listed$Year_Plan <- year(listed$Rec_Plan_Date)
# 
# no_plan <- filter(listed, is.na(listed$Year_Plan))
```
